(working title) Local Physiological Adaptation in \textit{Mimulus cardinalis}

### 2015-04-28 ms/ms.Rnw
Worked on Results.

As with previous studies in \textit{Mimulus} and many other species [CITES, Huey_etal, mosquitoes, what else?], latitude \textit{per se} predicted trait variation more than many obvious environmental variables (e.g. temperature or precipitation) that covary with latitude. We used Random Forest regression \citep{Liaw_Wiener_2002} to identify putative climatic factors underlying trait-latitude associations in \textit{M. cardinalis}. We did this by looking for overlap between climatic variables that best predict latitude of \textit{M. cadinalis} occurence records and climatic variables that best predict trait variation across our 16 focal populations. We used a representative set of \Sexpr{nrow(thinnedOccClimate)} recent (since 2000) known \textit{M. cardinalis} occurences. These occurences were thinned by 50\% to correct for uneven sampling from a comprehensive set of herbarium records and an exhaustive field survey in 2010-11 \citep{Angert_ENM}. For each occurence, we used a 90m digital elevation model from HydroSHEDS \citep{Lehner_etal_2006} to extract elevation. Monthly interpolated climate layers were calculated using ClimateWNA \citep{Wang_etal_2012}, which accurately downscales climate data specifically for the rugged topography of western North America. For each occurence, we calculated bioclimatic variables using the biovars function in the R package \pkg{dismo} \citep{Hijmans_etal_2014}. In total, we included \Sexpr{length(climVars) / 2} cliamte variables, 9 from ClimateWNA and 15 bioclimatic variables (Table~\ref{table:TableS_ClimVars}). For each variable, we calculated both a 30-year normal by averging annual values between 1981 and 2010 and 30-year coefficient of variation, a standardized metric of interannual climatic variation. Temperatures were converted to Kelvin to be on an interval scale appropriate for calculating the coefficient of variation.

Selection at given site might not reflect simply the local environment, but rather the average environment in a surrounding region, since alleles will experience selection across populations linked by migration. We therefore calculated spatially averaged \cdm{maybe spatially-integrated is better term?} climate means and coefficients of variation. We sampled climate at 1000 random points (at 90-m resolition) within a 10 km buffer around focal populations. Since \textit{M. cardinalis} is found exclusively in riparian areas, we only selected points along streams using the National Hydrogeoraphy Dataset \citep{NHD}. Climatic means and CVs were weighted by the predicted suitability of each point, using ENM models from \cite{Angert_ENM}.

To identify climatic variables that were closely associated with latitude in \textit{M. cardinalis}, we used a "Variable Selection Using Random Forest" (VSURF) algorithm in the R \pkg{VSURF} version 0.8.2 \citep{Genuer_etal_2014}. In this algorithm, climatic variables are regressed against latitude, and the most important variables were selected after a three-stage selection process that progressively excludes unimportant variables.

Predictions: looked for climatic variables (means or CVs) that were strong predictors of both population latitudes (from all card pops) and trait variation in 16 focal popultions. If spatial average environment is important, then expect greater predictive power and closer association with latitude-predictors than point averages.

Evaluated fit on test set (correlation between predicted versus actual latitude) and on pseudoabsence set.
, Despite this common, and long-standing pattern, only certain traits In particular, adaptation to the local abiotic environment is generally thought to be the major cause of natural selection. Physiology, the study of organismal function, connects fitness to the abiotic environment, yet we actually know little about the physiological basis of adaptation to different environments within a species. To address this question, we looked at physiological variation across a broad latitudual gradient within \textit{Mimulus cardinalis}...

% Methods for spatial-average selection
For 16 focal populations, I took the average climate within a 10 km buffer, clipped to streams based on the National Hydrological Database \citep{NHD} weighted by climatic suitability determined [multimodel average/ensemble?] from ecological niche modeling \citep{Angert_ENM}.

@@

\subsection*{Coordinated latitudinal clines in germination, growth, and photosynthesis}

@@

\subsection*{Latitude and phenotypic clines are associated with more variable precipitation}

@@

>>>>

As with previous studies in \textit{Mimulus} and many other species [CITES, Huey_etal, mosquitoes, what else?], latitude \textit{per se} predicted trait variation more than many obvious environmental variables (e.g. temperature or precipitation) that covary with latitude. We used Random Forest regression \citep{Liaw_Wiener_2002} to identify putative climatic factors underlying trait-latitude associations in \textit{M. cardinalis}. We did this by looking for overlap between climatic variables that best predict latitude of \textit{M. cadinalis} occurence records and climatic variables that best predict trait variation across our 16 focal populations. For brevity, we refer to these as Climate-Latitude and Climate-Trait variables. We selected Climate-Latitude and Climate-Trait variables independently using Random Forest (VSURF algorithm in the R \pkg{VSURF} version 0.8.2 \citep{Genuer_etal_2014}). From VSURF models, we kept only variables selected for prediction, the most stringent criteria.

For these analyses, we compiled a representative set of \Sexpr{nrow(thinnedOccClimate)} recent (since 2000) known \textit{M. cardinalis} occurences. These occurences were thinned by 50\% to correct for uneven sampling from a comprehensive set of herbarium records and an exhaustive field survey in 2010-11 \citep{Angert_ENM}. For each occurence, we used a 90m digital elevation model from HydroSHEDS \citep{Lehner_etal_2006} to extract elevation. Monthly interpolated climate layers were calculated using ClimateWNA \citep{Wang_etal_2012}, which accurately downscales climate data specifically for the rugged topography of western North America. For each occurence, we calculated bioclimatic variables using the biovars function in the R package \pkg{dismo} \citep{Hijmans_etal_2014}. In total, we included \Sexpr{length(climVars) / 2} cliamte variables, 9 from ClimateWNA and 15 bioclimatic variables (Table~\ref{table:TableS_ClimVars}). For each variable, we calculated both a 30-year normal by averging annual values between 1981 and 2010 and 30-year coefficient of variation, a standardized metric of interannual climatic variation. Temperatures were converted to Kelvin to be on a ratio scale appropriate for calculating the coefficient of variation.

@@

\subsection*{A coordinated latitudinal cline in germination, growth, and photosynthesis}

@@

\subsection*{Climatic variability best explains phenotypic divergence}

@@

\subsection*{Spatially-averaged climate}

% Methods for spatial-average selection
Selection at given site might not reflect simply the local environment, but rather the average environment in a surrounding region, since alleles will experience selection across populations linked by migration. We therefore calculated spatially averaged climate means and coefficients of variation. We sampled climate at 1000 random points (at 90-m resolition) within a 10 km buffer around focal populations. Since \textit{M. cardinalis} is found exclusively in riparian areas, we only selected points along streams using the National Hydrogeoraphy Dataset \citep{NHD}. Climatic means and CVs were weighted by their climatic suitability as determined using a multimodel ensemble average of ecological niche models \citep{Angert_ENM}.
